The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-07-10 manuscript.Rmd
edit mansucript, fix refs

Predicting the geographic range of species, and how they will be impacted by global change has undergone important advances recently [@xxxx]. However, it is generally recognized that to obtain reliable results such efforts require to have accurate geographic information for a minimum of 10 specimens but having at least 30 is preferred [@xxxx]. From the iDigBio data, we found that `r prop_10_specimens_species`% of species are known from less than 10 records and `r prop_30_specimens_species`% from less than 30 records. This suggests that our understanding of the geographic range of species is limited as most species are known from only a few locations, while efforts to estimate their current and future ranges based on occurrence data are rendered moot by the low number of data points available.

@@

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to efficiently document biodiversity, and build on our existing knowledge to collect information to improve our understanding of marine systems.

DNA sequences have been one of the first test case on how to link biodiversity data across repositories. These initial efforts have revealed the lack of infrastructure to use data formats designed for linked data, the lack of ontologies, and the fragility of the identifiers used [@Stucky2014; @Walls2014; @Guralnick2014]. As new types of data will be added and linked to existing repositories, the community needs to develop modern and robust ways to integrate heterogeneous biodiversity data sources [@Guralnick2015; @Deck2015].

Digitization efforts are catching up on already fragmented and incomplete records. Even if technological advances can greatly speed up the documentation of species and their distributions, taxonomists and experts are needed  to collect, sort, process, and identify samples. Taxonomic expertise within the US is dwindling, and there are now only a handful of US natural history museums with active marine invertebrates collections.

Crowd-sourced data and citizen science initiatives have been proposed to augment data about species abundance and distributions [@Pimm2014]. For marine taxa, beyond the application to near-shore habitats on easily recognizable species [@Delaney2008], large-scale efforts are currently not possible given the specialized equipment needed to sample most marine habitats and the expertise required to identify species. There are however opportunities for citizen science to help curate biodiversity data, while educating and engaging the public with marine conservation issues. For instance it has been successfully used to classify plankton images [@Robinson2017], or transcribe hand-written museum labels [@Hill2012].

>>>>

Predicting the geographic range of species, and how they will be impacted by global change has undergone important advances recently [@Guillera-Arroita2017], but remain scarecely used in the marine realm [@Robinson2011]. One limitation is the low number of observations available for each species: it is generally recognized that to obtain reliable results such efforts require to have accurate geographic information for a minimum of 10 specimens but having at least 30 is preferred [@Wisz2008]. From the iDigBio data, we found that `r prop_10_specimens_species`% of species are known from less than 10 records and `r prop_30_specimens_species`% from less than 30 records. This suggests that our understanding of the geographic range of species is limited as most species are known from only a few locations, while efforts to estimate their current and future ranges based on occurrence data are rendered moot by the low number of data points available.

@@

Given the urgency of the biodiversity crisis, we need to develop data-driven initiatives to document biodiversity more efficiently, and build on our existing knowledge to collect data that will improve our understanding of marine systems.

DNA sequences have been one of the first test case on how to link biodiversity data across repositories. These initial efforts have revealed the lack of infrastructure to use data formats specifically designed for this task, the lack of ontologies available for biodiversity data, and the fragility of the identifiers [@Stucky2014; @Walls2014; @Guralnick2014]. As new types of data will be collected, added, and linked across repositories, the community needs to develop modern and robust ways to integrate these heterogeneous biodiversity data sources [@Guralnick2015; @Deck2015].

Digitization efforts are catching up on already fragmented and incomplete records. Even if technological advances can greatly speed up the documentation of species and their distributions, taxonomists and experts are needed  to collect, sort, process, and identify samples. Taxonomic expertise within the US is dwindling, and there are now only a handful of US natural history museums with active marine invertebrates collections. This lack of resources is not sustainable in the face of the challenges and discoveries that marine ecosystems have to offer.

Crowd-sourced data and citizen science initiatives have been proposed to augment data about species abundance and distributions [@Pimm2014]. Beyond the application to near-shore habitats on easily recognizable species [@Delaney2008], large-scale efforts are currently not possible for marine ecosystems given the specialized equipment needed to sample most habitats and the expertise required to identify species. There are however opportunities for citizen science to help curate biodiversity data, while educating and engaging the public with marine conservation issues. For instance citizen science has been used successfully to classify plankton images [@Robinson2017], or transcribe hand-written museum labels [@Hill2012].
